{\fontfamily{ptm}\selectfont Linear Convergence of Stochastic Nesterov's Accelerated Proximal Gradient method under Interpolation Hypothesis, Truth or Just a Dream?}

### 2025-07-29 Notebook LaTEX/SAGD Under Interpolation/main.tex
>w<, small changes

\begin{abstract} 
    This file is for communication purposes between collaborators. 
    In brief, we think that the conditions required for linear convergence rate of Stochastic Nesterov's accelerated gradient (or proximal gradient) is too precarious to hold, even with interpolation hypothesis. 
    Instead of attacking the problem head on, this file will characterize the conditions required for linear convergence rate. 
    We place specific constraints on the random variable representing the error made when estimating gradient via some random variables. 
\end{abstract}

@@

        \textbf{We very much would like $\epsilon$ to be a constant to make the algebra possible when deriving the convergence rate of the algorithm. }

@@

        The following lemma gives a proximal gradient inequality when $\tilde \nabla f(x)$ is an estimate by some random variable, \textbf{and it is the precursor.} 

@@

@@

@@

@@

@@

                \item it is initialized with $v_{-1} = x_{-1}$, $\alpha_0 = 1$ and, the momentum sequence $(\alpha_k)_{k \ge 0}$ satisfies Definition \ref{def:relax-momen-seq}. 

@@

@@

                \item [(b)] We assumed that the sequence $(\alpha_k)_{k \ge 0}$ satisfies\\ for all $k \ge 1$: $\rho_{k - 1}(1 - \alpha_{k})\alpha_{k - 1}^2 = \alpha_{k}(\alpha_{k} - \mu/L)$. 

@@

            For \circled{3}, we used Lemma \ref{lemma:seq-properties} (i) and (b) because the sequence $\alpha_k \le 1$, therefore $1 - \alpha_k \le 0$. 
            Next, $\alpha_0 = 1 > \mu/L$, Lemma \ref{lemma:seq-properties} (ii) applies too, hence it also has $L\alpha_k - \mu \ge 0$. 
            Since $L > \mu$ always, the coefficient on $\Vert x_{k - 1} - v_{k - 1}\Vert^2$ is always $\le 0$, therefore the term can be removed to give an inequality. 

@@

        \begin{theorem}[SNAG convergence rate]\label{thm:snag-convergence}

@@

            Assuming that it's generated by the stochastic gradient operator as in Definition \ref{def:stoch-pg-opt-rel-err}. 

@@

                \right)\expect_k \Phi_0 + 
                \sum_{j = 0}^{k}\left(

@@

            Since this is true for all $k \ge 0$ in $\N$, we are now in position to verify the following inductive hypothesis 

@@

                \right)\expect_k \Phi_0 + 
                \sum_{j = 0}^{k}\left(

@@

                    \right)\expect_k \Phi_0 

@@

                \right)\expect_k \Phi_0
                + \sum_{j = 1}^{k + 1}\left(

@@

                \right)\expect_0 R_{k + 1 - j}

@@

                \right)\expect_k \Phi_0
                + \sum_{j = 0}^{k + 1}\left(

@@

    \subsection{Discussion of he error schedules}
        Let's talk about Theorem \ref{thm:snag-convergence} here. 

@@

            \item $\Phi_0$, the base case and a bounded quantities, depends on initial iterate $v_{-1}, x_{-1}$. 
            \item $\lambda_k$, a sequence controling the convergence, involving relaxation parameters $\rho_{k - 1}$ and, the error $\epsilon_k$. 
            \item $R_k$, an error term that, it's none trivial to group with the other terms of the inequality. 

@@

        Under some assumptions, and conditions, we can have some type of convergence rate. 
\section{Interpolation helps but, it's very unclear if it's sufficient}
    \subsection{The no error case}
    \subsection{The degenerate case}

>>>>

\begin{abstract} 
    This file is for communication purposes between collaborators. 
    In brief, we think that the conditions required for fast linear convergence rate, with a square root on the condition number of Stochastic Nesterov's accelerated gradient (or proximal gradient) is too precarious to hold, even with interpolation hypothesis. 
    Instead of attacking the problem head on, this file will characterize the conditions required for fast linear convergence rate. 
    We place specific constraints on the random variable representing the error made when estimating gradient via some random variables. 
\end{abstract}

@@

        \textbf{We very much would like $\epsilon$ to be a constant instead of a random variable to make the algebra possible when deriving the convergence rate of the algorithm. }

@@

        The following lemma gives a proximal gradient inequality when $\tilde \nabla f(x)$ is an estimate by some random variable, \textbf{and it is the precursor that gives the analysis of our convergence rate.} 

@@

    \subsection{discussion}

@@

        \begin{align}\label{eqn:stoch-grad-epsilon}

@@

        \textbf{Ok, what about the relation between this error term \eqref{eqn:stoch-grad-epsilon} and the strong growth conditions that is usually found in the literature?}
        There is no direct equality relationship between the Strong Growth condition of the stochastic gradient which is usually used in the literature, with the condition we have here. 
        Nonetheless, an indirect inequality relation exists. 
        Suppose that the gradient of $f$ satisfies strong growth condition with constant $\rho$, so it has $\expect \left\Vert \tilde \nabla f(x)\right\Vert^2 \le \rho \Vert \nabla f(x)\Vert^2$, then $\left\Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert$ squared has: 
            & \left(
                \expect \left \Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert
            \expect \left \Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert^2 
            &= \expect \left\Vert \nabla f(x) - \expect \tilde \nabla f(x) \right\Vert^2 
            &= \expect \left\Vert \tilde \nabla f(x)\right\Vert^2 
            - \Vert \nabla f(x)\Vert^2 
            \le (\rho - 1)\Vert \nabla f(x)\Vert^2. 
        The last two equalities, we used the $\expect (X - \expect X)^2 = \expect X^2 - (\expect X)^2$ with $X$ being each of the element of the vector and, the linearity of expectations to apply it individually and them sum it up. 
        The first inequality also uses that equality but $X = \left\Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert$ instead, and $\expect (X - \expect X)^2 \ge 0$ always. 
        The last inequality comes from the assumed strong-growth conditions. 
        Strong growth condition does make $\epsilon$ smaller if $x$ is close to the set of minimizer and this is the result: 
            \expect \left\Vert
            \right\Vert &\le 
            \sqrt{\rho - 1} \Vert \nabla f(x)\Vert. 
        Keeping the strong growth condition, let's assume that $\tilde x$ is generated by stochastic gradient, so that the proximal operator has $g \equiv 0$, then $\tilde x = x - L^{-1}\tilde \nabla f (x)$ directly. 
        Assuming that relation $m \le \Vert z - \tilde x\Vert \le M$ exists then \eqref{eqn:stoch-grad-epsilon} has upper bound
            & \frac{M}{m} \frac{\expect\left\Vert \nabla f(x) - \tilde \nabla f(x)\right\Vert}{\expect \Vert x - \tilde x\Vert}
            &\le \frac{M}{m}\frac{\sqrt{\rho - 1}\Vert \nabla f(x)\Vert}{L^{-1}\expect \left\Vert\tilde \nabla f(x)\right\Vert}
            &\le \frac{LM}{m}
            \frac{\sqrt{\rho - 1}\Vert \nabla f(x)\Vert}{
                \min_{y \in \Omega(x)}\left\{
                    \left\Vert y \right\Vert: y\neq \mathbf 0
                \right\}\prob\left(\left\Vert\tilde \nabla f(x)\right\Vert \neq 0\right)
        \todoinline{NOT FINISHED YET. It seems like no pretty relaton exists. }

@@

    The sequence is powerful, instead of an equality relation, the sequence can instead satisfies $\alpha_k \in (\mu/L, 1)$. 

@@

                \item it is initialized with $v_{-1} = x_{-1}$, $\alpha_0 = 1$ and, the momentum sequence $(\alpha_k)_{k \ge 0}$ is given as in Lemma \ref{lemma:seq-properties}. 

@@

                + \frac{\epsilon_0}{2}\expect_0\left[\Vert v_{-1} - x_0\Vert^2\right]. 

@@

                \item [(b)] We assumed that the sequence $(\alpha_k)_{k \ge 0}$ satisfies\\ for all $k \ge 1$: $\rho_{k - 1}(1 - \alpha_{k})\alpha_{k - 1}^2 = \alpha_{k}(\alpha_{k} - \mu/L)$, hence Lemma \ref{lemma:seq-properties} applies. 

@@

            For \circled{2}, we used Lemma \ref{lemma:seq-properties} (i) because of what is assumed in (b) so, the sequence has $\alpha_k \le 1$, therefore $1 - \alpha_k \le 0$. 
            Next, $\alpha_0 = 1 > \mu/L$, Lemma \ref{lemma:seq-properties} (ii) applies too, hence $L\alpha_k - \mu \ge 0$. 
            Since $L > \mu$ always, the coefficient on $\Vert x_{k - 1} - v_{k - 1}\Vert^2$ is always $\le 0$, therefore the term can be removed to keep the inequality. 

@@

        \begin{theorem}[SNAG generic convergence rate]\label{thm:snag-generic-cnvg}

@@

            Assume that the momentum sequence $(\alpha_k)_{k \ge 0}$, stochastic gradient $\tilde \nabla f$, and $v_{-1}, x_{-1}$ as in Theorem \ref{thm:snag-descent}. 

@@

                \right)\expect \Phi_0 + 
                \sum_{j = 0}^{k - 1}\left(

@@

            Let $k \ge 1$, We are now in position to verify the following inductive hypothesis: 

@@

                \right)\expect_0 \Phi_0 + 
                \sum_{j = 0}^{k - 1}\left(

@@

                    \right)\expect_0 \Phi_0 
                    + \sum_{j = 0}^{k - 1}\left(

@@

                \right)\expect_0 \Phi_0
                + \sum_{j = 1}^{k}\left(

@@

                \right)\expect_0 R_{k - j}

@@

                \right)\expect_0 \Phi_0

@@

    \subsection{Discussion of error schedules}
        Let's talk about Theorem \ref{thm:snag-generic-cnvg} here. 

@@

            \item $\Phi_0$, the base case and a bounded quantity, depends on initial iterate $v_{-1}, x_{-1}$. 
            \item $\lambda_k$, a sequence controlling the convergence, involving relaxation parameters $\rho_{k - 1}$ and, the error $\epsilon_k$. 
            \item $R_k$, an error term related to $\epsilon_k$ and, $\Vert y_k - x_k\Vert^2$, it's none trivial to group with the other terms. 

@@

        Without necessarily introducing the interpolation hypothesis, we have linear ``convergence" with an extra additive term under mild assumption. 
        The following propositions simply stated the sufficient conditions on $\Phi_k, \lambda_k, R_k$ for the inequality to yield a linear convergence rate with additive error, nothing more. 
        \begin{proposition}[SNAG kinda converges linearly]\;\label{prop:snag-kinda-converge}\\
            Suppose that the sequence $(y_k, x_k, v_k)_{k \ge 0}$ is generated by algorithms satisfiying \ref{def:SNAG}. 
            Suppose that that $\Phi_k, \lambda_k, R_k$ are as given in Theorem \ref{thm:snag-generic-cnvg}. 
            Define
                \item $\Lambda = \sup_{i \in \N} \lambda_i$. 
                \item An upper bound $\bar R \ge \expect R_k$ exists for all $k \ge 0$. 
                \item An upper bound $\overline \epsilon = \sup_{i \in \N \cup\{0\}} \epsilon_i$ exists. 
            If in addition, we have $\overline \epsilon < \sqrt{\mu L}$, then it satisfies for all $k \ge 1$ the following inequality: 
                \expect \Phi_k \le \Lambda^k \expect \Phi_0 + \frac{\overline R}{1 - \Lambda}. 
            Where 
            \Lambda = \left(
                1 - \sqrt{\frac{\mu}{L}}
                    \frac{L}{L - \overline \epsilon}
        \end{proposition}
            We can prove it with the following intermediate results: 
                \item[(a)] For all $k \ge 1$, $\alpha_k = \sqrt{\mu/L}$ satisfies recurrences $(1 - \alpha_k)\alpha_{k - 1}^2 = \alpha_k(\alpha_k - \mu/L)$ where $\rho_k = 1$ for all $k \ge 1$, and $\rho_0 = \mu/L$. 
                \item[(b)] When $(\epsilon_k)_{k \ge 0}$ satisfies $\epsilon_k \le  \bar \epsilon = \sup_{i \in \N\cup\{0\}} \epsilon_i < \sqrt{\mu L}$, it has $\lambda_k \in (0, 1)$ for all $k \ge 1$, with $\Lambda = \sup_{i\in\N} \lambda_i < 1$. 
            These results will be proved at the end. 
            Consider starting with the inequality proved in Theorem \ref{thm:snag-descent}: 
                \expect \Phi_k 
                &\le \left(
                \right) \expect \Phi_0
                + \sum_{j = 1}^{k - 1}\left(
                    \prod_{j = k - j}^{k - 1} \lambda_{i}
                + \expect R_k
                &\underset{\circled{1}}{\le} \Lambda^k \expect \Phi_0 + \sum_{j = 0}^{k - 1} \Lambda^{j + 1} \expect R_{k - j - 1} 
                + \expect R_k
                &= \Lambda^k \expect \Phi_0 + \sum_{j = 1}^{k} \Lambda^{j} \expect R_{k - j} 
                + \expect R_k
                &= \Lambda^k \expect \Phi_0 + \sum_{j = 0}^{k} \Lambda^{j} \expect R_{k - j} 
                &\underset{\circled{2}}{\le} \Lambda^k \expect \Phi_0 + \overline R \sum_{j = 0}^{k} \Lambda^{j} 
                &\underset{\circled{3}}{\le} \Lambda^k \expect \Phi_0 + \frac{\overline R}{1 - \Lambda}. 
            At \circled{1} we used that $\Lambda \ge \lambda_k$. 
            At \circled{2} we used that $\overline R \ge \expect R_{k}$ for all $k \ge 0$. 
            At \circled{3} we used the results from (b) that $\lambda_k \le \Lambda$ and $\Lambda < 1$, hence it has the convergence of the geometric series. 
            \par \textbf{Proof of (a)}
            When $L > \mu > 0$, so the objective $F$ is strongly convex, the usual $(1 - \alpha_k)\alpha_{k - 1}^2 = \alpha_k(\alpha_k - \mu/L)$ recurrence relations admits a constant solution $
            \alpha_k = \sqrt{\mu/L}$ for all $k \ge 1$. 
            As presented in Theorem \ref{thm:snag-descent}, the momentum sequence $(\alpha_k)_{k \ge 0}$ has a relaxation parameter $(\rho_k)_{k \ge 1}$. 
            In this case, it means that $\rho_{k - 1} = 1$ for all $k \ge 1$. 
            Since when $k = 0$, $\alpha_0 = 1$, it makes an exception on $\rho_0$, which is: 
                \rho_0 &= \frac{\sqrt{\mu/L}\left(\sqrt{\mu/L} - \mu/L\right)}{1 - \sqrt{\mu/L}} = \frac{\mu}{L}. 
            \par \textbf{Proof of (b)}.
            If $\sup_{k \in \N \cup \{0\}}\epsilon_{k} < \sqrt{\mu L}$ for all $k \ge 0$, by the definition of $\lambda_k$, it has: 
                & \sup_{k \in \N}\lambda_k
                &= \sup_{k \in \N} \left(
                    1 - \sqrt{\frac{\mu}{L}}
                    1 - \sqrt{\frac{\mu}{L}}
                    \frac{L}{L - \sup_{k \in \N}\epsilon_{k - 1}}
                &< \left(
                    1 - \sqrt{\frac{\mu}{L}}
                    \frac{L}{1 - \sqrt{\mu L}}
                \right) = 1. 
            At \circled{1}, we used the fact that $\rho_k = 1$ for all $k \ge 1$ and,$\rho_k = \mu/L$ for $k = 0$, therefore, it $(\forall k \ge 0)\; \rho_k \le 1$. 
            So $\Lambda < 1$.

\section{Interpolation helps but, it's very unclear if it's sufficient}
    Under the frameworks of our analysis, the interpolation assumption helps with controlling the errors, but it's unclear to what extent, it will help us to obtain a convergence rate faster than what is already in the literatures. 
    \subsection{The no error case}
    \subsection{The degenerate case}
